---
title: "Mini Data-Analysis Deliverable 1"
output: html_document
---

# Welcome to your (maybe) first-ever data analysis project!

And hopefully the first of many. Let's get started:

1.  Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package by typing the following into your **R terminal**:

<!-- -->

    install.packages("devtools")
    devtools::install_github("UBC-MDS/datateachr")

2.  Load the packages below.

```{r}
library(datateachr)
library(tidyverse)
```

3.  Make a repository in the <https://github.com/stat545ubc-2023> Organization. You can do this by following the steps found on canvas in the entry called [MDA: Create a repository](https://canvas.ubc.ca/courses/126199/pages/mda-create-a-repository). One completed, your repository should automatically be listed as part of the stat545ubc-2023 Organization.

# Instructions

## For Both Milestones

-   Each milestone has explicit tasks. Tasks that are more challenging will often be allocated more points.

-  Each milestone will be also graded for reproducibility, cleanliness, and coherence of the overall Github submission. 

- While the two milestones will be submitted as independent deliverables, the analysis itself is a continuum - think of it as two chapters to a story. Each chapter, or in this case, portion of your analysis, should be easily followed through by someone unfamiliar with the content. [Here](https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/) is a good resource for what constitutes "good code". Learning good coding practices early in your career will save you hassle later on!

- The milestones will be equally weighted. 

## For Milestone 1

**To complete this milestone**, edit [this very `.Rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-1.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work below --->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to the mini-analysis GitHub repository you made earlier, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 36 points: 30 for your analysis, and 6 for overall reproducibility, cleanliness, and coherence of the Github submission. 

# Learning Objectives

By the end of this milestone, you should:

-   Become familiar with your dataset of choosing
-   Select 4 questions that you would like to answer with your data
-   Generate a reproducible and clear report using R Markdown
-   Become familiar with manipulating and summarizing your data in tibbles using `dplyr`, with a research question in mind.

# Task 1: Choose your favorite dataset 

The `datateachr` package by Hayley Boyce and Jordan Bourak currently composed of 7 semi-tidy datasets for educational purposes. Here is a brief description of each dataset:

-   *apt_buildings*: Acquired courtesy of The City of Toronto's Open Data Portal. It currently has 3455 rows and 37 columns.

-   *building_permits*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 20680 rows and 14 columns.

-   *cancer_sample*: Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.

-   *flow_sample*: Acquired courtesy of The Government of Canada's Historical Hydrometric Database. It currently has 218 rows and 7 columns.

-   *parking_meters*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 10032 rows and 22 columns.

-   *steam_games*: Acquired courtesy of Kaggle. It currently has 40833 rows and 21 columns.

-   *vancouver_trees*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 146611 rows and 20 columns.

**Things to keep in mind**

-   We hope that this project will serve as practice for carrying our your own *independent* data analysis. Remember to comment your code, be explicit about what you are doing, and write notes in this markdown document when you feel that context is required. As you advance in the project, prompts and hints to do this will be diminished - it'll be up to you!

-   Before choosing a dataset, you should always keep in mind **your goal**, or in other ways, *what you wish to achieve with this data*. This mini data-analysis project focuses on *data wrangling*, *tidying*, and *visualization*. In short, it's a way for you to get your feet wet with exploring data on your own.

And that is exactly the first thing that you will do!

1.1 **(1 point)** Out of the 7 datasets available in the `datateachr` package, choose **4** that appeal to you based on their description. Write your choices below:

**Note**: We encourage you to use the ones in the `datateachr` package, but if you have a dataset that you'd really like to use, you can include it here. But, please check with a member of the teaching team to see whether the dataset is of appropriate complexity. Also, include a **brief** description of the dataset here to help the teaching team understand your data.

<!-------------------------- Start your work below ---------------------------->

1: CHOICE_1\_vancouver_trees\
2: CHOICE_2\_parking_meters\
3: CHOICE_3\_flow_sample\
4: CHOICE_4\_building_permits

<!----------------------------------------------------------------------------->

1.2 **(6 points)** One way to narrowing down your selection is to *explore* the datasets. Use your knowledge of dplyr to find out at least *3* attributes about each of these datasets (an attribute is something such as number of rows, variables, class type...). The goal here is to have an idea of *what the data looks like*.

*Hint:* This is one of those times when you should think about the cleanliness of your analysis. I added a single code chunk for you below, but do you want to use more than one? Would you like to write more comments outside of the code chunk?

<!-------------------------- Start your work below ---------------------------->
**Dataset exploration**  
To explore attributes of each of the four datasets, I will use the glimpse() function from the *dplyr* package to find out the number of columns (variables), rows (observations), and the class type for each variable. I will check the class of the dataset using the class() function in base R. Finally, I will look up the R documentation associated with each dataset to read the  description of the dataset and its' variables.

```{r}
#Vancouver Street Trees
glimpse(vancouver_trees)
class(vancouver_trees)
?vancouver_trees
```
Using the glimpse() function, the dataset contains 146,611 rows and 20 variables. 7 variables are double precision (continuous quantitative value), 12 are character strings, and one is a date. The dataset is a tibble. The information it contains is about public trees in the City of Vancouver. It is updated weekly, except for tree attributes which are rarely updated.


```{r}
#Vancouver Parking Meters
glimpse(parking_meters)
class(parking_meters)
?parking_meters
```
This dataset contains 10,032 rows and 22 variables. Two of the variables are double precision and 20 are character strings. The dataset is a tibble. The information it contains is about the rates and time limits for parking meters in Vancouver, shown for entire block faces rather than individual meters. The data is updated weekly.

```{r}
#Annual Bow River Banff Flow Rate Extremes
glimpse(flow_sample)
class(flow_sample)
?flow_sample
```
This dataset contains 218 rows and 7 variables. Four of the variables are double precision and three are character strings. The dataset is a tibble. The information contained in this dataset is historical flow rate extrema data collected by the Government of Canada at station 05BB001, located on the Bow River near Banff, Alberta.

```{r}
#Vancouver Building Permits
glimpse(building_permits)
class(building_permits)
?building_permits
```
This dataset contains 20,680 rows and 14 variables. Three of the varaibles are double precision, one is a date, and 10 are character strings. The dataset is a subclass of tibble. The information it contains is about building permits issued by the City of Vancouver since the year 2017, updated monthly for the current year.
<!----------------------------------------------------------------------------->

1.3 **(1 point)** Now that you've explored the 4 datasets that you were initially most interested in, let's narrow it down to 1. What lead you to choose this one? Briefly explain your choice below.

<!-------------------------- Start your work below ---------------------------->
I will use the vancouver_trees datatset for my analysis. One of the things I appreciate about Vancouver is the number and variety of trees that have been planted throughout the city, and this dataset contains a a large number of qualitative and quantitative variables relating to both attributes of the trees themselves (species, height, diameter, etc) as well as their location in geographic space that I am interested in exploring. 
<!----------------------------------------------------------------------------->

1.4 **(2 points)** Time for a  final decision! Going back to the beginning, it's important to have an *end goal* in mind. For example, if I had chosen the `titanic` dataset for my project, I might've wanted to explore the relationship between survival and other variables. Try to think of 1 research question that you would want to answer with your dataset. Note it down below.

<!-------------------------- Start your work below ---------------------------->
I am interested in exploring the relationship between the tree species and date planted in this dataset, specifically how did the most commonly planted tree species change each decade?

<!----------------------------------------------------------------------------->

# Important note

Read Tasks 2 and 3 *fully* before starting to complete either of them. Probably also a good point to grab a coffee to get ready for the fun part!

This project is semi-guided, but meant to be *independent*. For this reason, you will complete tasks 2 and 3 below (under the **START HERE** mark) as if you were writing your own exploratory data analysis report, and this guidance never existed! Feel free to add a brief introduction section to your project, format the document with markdown syntax as you deem appropriate, and structure the analysis as you deem appropriate. If you feel lost, you can find a sample data analysis [here](https://www.kaggle.com/headsortails/tidy-titarnic) to have a better idea. However, bear in mind that it is **just an example** and you will not be required to have that level of complexity in your project.

# Task 2: Exploring your dataset 

If we rewind and go back to the learning objectives, you'll see that by the end of this deliverable, you should have formulated *4* research questions about your data that you may want to answer during your project. However, it may be handy to do some more exploration on your dataset of choice before creating these questions - by looking at the data, you may get more ideas. **Before you start this task, read all instructions carefully until you reach START HERE under Task 3**.

2.1 **(12 points)** Complete *4 out of the following 8 exercises* to dive deeper into your data. All datasets are different and therefore, not all of these tasks may make sense for your data - which is why you should only answer *4*. 

Make sure that you're using dplyr and ggplot2 rather than base R for this task. Outside of this project, you may find that you prefer using base R functions for certain tasks, and that's just fine! But part of this project is for you to practice the tools we learned in class, which is dplyr and ggplot2.

1.  Plot the distribution of a numeric variable.
2.  Create a new variable based on other variables in your data (only if it makes sense)
3.  Investigate how many missing values there are per variable. Can you find a way to plot this?
4.  Explore the relationship between 2 variables in a plot.
5.  Filter observations in your data according to your own criteria. Think of what you'd like to explore - again, if this was the `titanic` dataset, I may want to narrow my search down to passengers born in a particular year...
6.  Use a boxplot to look at the frequency of different observations within a single variable. You can do this for more than one variable if you wish!
7.  Make a new tibble with a subset of your data, with variables and observations that you are interested in exploring.
8.  Use a density plot to explore any of your variables (that are suitable for this type of plot).

2.2 **(4 points)** For each of the 4 exercises that you complete, provide a *brief explanation* of why you chose that exercise in relation to your data (in other words, why does it make sense to do that?), and sufficient comments for a reader to understand your reasoning and code.

<!-------------------------- Start your work below ---------------------------->
#Public Street Trees in the City of Vancouver  
##Introduction
This is an exploratory data analysis using the R *tidyverse* packages, *dplyr* and *ggplot2*, to investigate a street tree dataset for the City of Vancouver in British Columbia, Canada. The open data portal for this dataset can be found [here.](https://opendata.vancouver.ca/explore/dataset/street-trees/information/?disjunctive.species_name&disjunctive.common_name&disjunctive.on_street&disjunctive.neighbourhood_name&location=6,50.02186,-122.91504)  
From the website, >The street tree dataset includes a listing of public trees on boulevards in the City of Vancouver and provides data on tree coordinates, species and other related characteristics. Park trees and private trees are not included in the inventory.  
In this analysis, I used *dplyr* for data manipulation and *ggplot2* for visualization. I used some basic functions and created simple plots to explore the dataset. Using what I learned from this initial data exploration, I developed research questions to investigate further.

##Data Overview  

To begin, I loaded the necessary packages using the library() function.
```{r}
library(datateachr)
library(tidyverse)
```

To get a quick overview of the dataset, I used the glimpse() and summary() functions.
```{r}
glimpse(vancouver_trees)
```

```{r}
summary(vancouver_trees)
```

From the glimpse() function I learned that this dataset contains 20 variables and 146,611 observations. 7 variables are double precision (continuous quantitative value), 12 are character strings, and one is a date.  
From the summary() function I learned information about the class of each variable and which variables have missing information (recorded as NA's - in this case, date_planted, longitude and latitude contain missing values).  

##Data Exploration  
Before beginning to look for relationships and patterns in the data, I created an object I named 'trees' from the dataset to be sure I did not alter the original dataset.

```{r}
trees <- vancouver_trees
```

**Distribution of a numeric variable**  
To begin, I plotted the height distribution of the trees in Vancouver to get an idea of the trend in height. The height value, stored in the variable height_range_id, is a numeric variable from 0 to 10 representing discrete height ranges for every 10 feet, where 0= 0-10 ft, 1=10-20 ft, 2=20-30ft, and so on.  
The height variable contains discrete numeric values, therefore I used the geom_bar() function (used to plot one discrete variable) to create a bar chart representing the height distribution for all of the trees in the dataset. The *ggplot2* approach to visualizing data uses a data input, the ggplot() function to create a plot, geom() functions to indicate the plotting style, and aesthetic specifications represented by aes() to map visual cues to variables.
```{r}
#Bar chart showing distribution of tree height
trees %>% ggplot(aes(height_range_id)) + 
 geom_bar(fill="palegreen3")+
  labs(title="Distribution of tree heights in Vancouver", x="Height range", y="Number of trees")+
  scale_x_binned()+
  scale_y_continuous(label=scales::comma)+
  theme_bw()
```
In the ggplot aesthetic mapping, I specified the x variable height_range_id. I binned the x axis using scale_x_binned() so that the axis values on the chart would correspond with the height range values. I created a title and x and y axis labels using the labs() function. With the scale_y_continuous() function, I added a comma to the y axis values which represent thousands of trees, so they would be easier to read.

From the bar chart we can see that most trees planted in Vancouver fall within the 1 and 2 height range bins, meaning that most trees are 30 ft or less in height.  


**Relationship between two variables**
The dataset contains information about tree diameter and whether or not the tree was planted with a root barrier. I examined the relationship between the continuous numeric variable, diameter, and the categorical variable, root_barrier, using boxplots. I was interested in how the presence of a root barrier might be related to tree size. In the ggplot aesthetic mapping I specified fill to be the root_barrier variable in order to color the two boxplots by the root_barrier output (Y or N). I added a logarithmic y axis using the scale_y_log10() function because the presence of outliers in the diameter values skewed the chart and made it difficult to compare the two boxplots. I used the scale_fill_manual() function to add specific colors to the boxplots and rename the legend title and values to "Root Barrier", "Yes" and "No".
```{r}
#Boxplots of tree diameter stratified by root barrier presence
trees %>% ggplot(aes(root_barrier, diameter, fill=root_barrier)) +
  geom_boxplot()+
  scale_y_log10()+
  scale_fill_manual(values=c("cornflowerblue", "darkorange"), name="Root Barrier", labels=c("No", "Yes"))+
  xlab("Root barrier")+
  ylab("Tree diameter, log10 transformed")+
  ggtitle("Tree diameter, stratified by presence of a root barrier")+
  theme_classic()
```
In this chart, it is clear that trees without a root barrier are larger on average than trees with a root barrier.

**Density Plot**  
Each tree that was planted has an associated planting date (except for those with missing date values, of which there are 76,548); therefore, a density plot of this numeric variable can be used to explore temporal patterns in the dataset. To visualize trends in planting over time, I created a density plot of the date_planted variable.  I used the function after_stat() and specified 'scaled' to scale the density estimate to a maximum of 1.

```{r}
#Density plot of planting date
trees %>% ggplot(aes(date_planted, after_stat(scaled)))+
   geom_density()+
  xlab("Date Planted")+
  ylab(NULL)+
  ggtitle("Density plot of planting date, scaled to 1")+
  theme_light()
```

**Filtering by a Variable**
Urban trees provide services like creating shade in the summer and filtering airborne pollutants. Red maple, silver linden, and American elm are three trees that are suitable for urban environments and are efficient at removing airborne particulate matter (PM2.5). To find out how many trees with these common names (i.e. not including cultivars) have been planted in Vancouver, I used the filter() function to select the three trees using the common_name variable. I made a bar chart and flipped the x and y axes using the coord_flip() function so the neighbourhood names would be listed on the y axis. I used the facet_wrap() function to make three comparison plots by common name, to look at patterns between tree type by neighbourhoods.
```{r}
#Bar chart of 3 air pollution-reducing trees by neighbourhood
trees %>% filter(common_name %in% c("RED MAPLE", "AMERICAN ELM", "SILVER LINDEN")) %>%
  ggplot(aes(neighbourhood_name, fill=common_name))+
  geom_bar()+
  coord_flip()+
  labs(title="Number of trees by common name and neighbourhood ", y="Number of trees", x=NULL)+
  facet_wrap(~common_name)+
  theme_bw()+
  theme(legend.position="none")
```
<!----------------------------------------------------------------------------->

# Task 3: Choose research questions 

**(4 points)** So far, you have chosen a dataset and gotten familiar with it through exploring the data. You have also brainstormed one research question that interested you (Task 1.4).  Now it's time to pick 4 research questions that you would like to explore in Milestone 2! Write the 4 questions and any additional comments below. 

<!--- *****START HERE***** --->
**Research Questions**
1.What are the 10 most commonly planted tree species in Vancouver, and what proportion of each species was planted with a root barrier?
2.What species of trees were planted between 1995-2000, and how do they compare to those planted between 2010-2015?
3.Where are the 10 tallest trees located in Vancouver, and what species cultivar are they?
4.What is the distribution of coniferous and deciduous trees by neighbourhood in Vancouver?


<!----------------------------->

# Overall reproducibility/Cleanliness/Coherence Checklist 

## Coherence (0.5 points)

The document should read sensibly from top to bottom, with no major continuity errors. An example of a major continuity error is having a data set listed for Task 3 that is not part of one of the data sets listed in Task 1. 

## Error-free code (3 points) 

For full marks, all code in the document should run without error. 1 point deduction if most code runs without error, and 2 points deduction if more than 50% of the code throws an error. 

## Main README (1 point)

There should be a file named `README.md` at the top level of your repository. Its contents should automatically appear when you visit the repository on GitHub.

Minimum contents of the README file:

-   In a sentence or two, explains what this repository is, so that future-you or someone else stumbling on your repository can be oriented to the repository.
-   In a sentence or two (or more??), briefly explains how to engage with the repository. You can assume the person reading knows the material from STAT 545A. Basically, if a visitor to your repository wants to explore your project, what should they know?

Once you get in the habit of making README files, and seeing more README files in other projects, you'll wonder how you ever got by without them! They are tremendously helpful.

## Output (1 point)

All output is readable, recent and relevant:

-   All Rmd files have been `knit`ted to their output md files. 
-   All knitted md files are viewable without errors on Github. Examples of errors: Missing plots, "Sorry about that, but we can't show files that are this big right now" messages, error messages from broken R code
-   All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
-   There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted.

(0.5 point deduction if any of the above criteria are not met. 1 point deduction if most or all of the above criteria are not met.)

Our recommendation: right before submission, delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant. Then, after your final commit and push to Github, CHECK on Github to make sure that everything looks the way you intended!

## Tagged release (0.5 points)

You've tagged a release for Milestone 1. 

### Attribution

Thanks to Icíar Fernández Boyano for mostly putting this together, and Vincenzo Coia for launching.
