---
title: "Mini Data Analysis Milestone 2"
output: github_document
---

*To complete this milestone, you can either edit [this `.rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are commented out with `<!--- start your work here--->`. When you are done, make sure to knit to an `.md` file by changing the output in the YAML header to `github_document`, before submitting a tagged release on canvas.*

# Welcome to the rest of your mini data analysis project!

In Milestone 1, you explored your data. and came up with research questions. This time, we will finish up our mini data analysis and obtain results for your data by: 

- Making summary tables and graphs 
- Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

We will also explore more in depth the concept of *tidy data.* 

**NOTE**: The main purpose of the mini data analysis is to integrate what you learn in class in an analysis. Although each milestone provides a framework for you to conduct your analysis, it's possible that you might find the instructions too rigid for your data set. If this is the case, you may deviate from the instructions -- just make sure you're demonstrating a wide range of tools and techniques taught in this class.

# Instructions

**To complete this milestone**, edit [this very `.Rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work here--->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to your mini-analysis GitHub repository, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 50 points: 45 for your analysis, and 5 for overall reproducibility, cleanliness, and coherence of the Github submission. 

**Research Questions**: In Milestone 1, you chose two research questions to focus on. Wherever realistic, your work in this milestone should relate to these research questions whenever we ask for justification behind your work. In the case that some tasks in this milestone don't align well with one of your research questions, feel free to discuss your results in the context of a different research question.

# Learning Objectives

By the end of this milestone, you should:

-   Understand what *tidy* data is, and how to create it using `tidyr`.
-   Generate a reproducible and clear report using R Markdown.
-   Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

# Setup

Begin by loading your data and the tidyverse package below:

```{r, message = FALSE}
library(datateachr) # <- might contain the data you picked!
library(tidyverse)
```

# Task 1: Process and summarize your data 

From milestone 1, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various data manipulation functions. 

### 1.1 (1 point) 

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

<!-------------------------- Start your work below ---------------------------->
1. What are the 10 most commonly planted tree species in Vancouver, and what proportion of each species was planted with a root barrier?
2. What species of trees were planted between 1995-2000, and how do they compare to those planted between 2010-2015?
3. Where are the 10 tallest trees located in Vancouver, and what species are they?
4. What is the distribution of coniferous and deciduous trees by neighbourhood in Vancouver?
<!----------------------------------------------------------------------------->

Here, we will investigate your data using various data manipulation and graphing functions.

### 1.2 (8 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question.

Ensure that the output of each operation is printed!

Also make sure that you're using dplyr and ggplot2 rather than base R. Outside of this project, you may find that you prefer using base R functions for certain tasks, and that's just fine! But part of this project is for you to practice the tools we learned in class, which is dplyr and ggplot2.

**Summarizing:**

1.  Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2.  Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3.  Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Compute the proportion and counts in each category of one categorical variable across the groups of another categorical variable from your data. Do not use the function `table()`!

**Graphing:**

6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.
7. Make a graph where it makes sense to customize the alpha transparency.

Using variables and/or tables you made in one of the "Summarizing" tasks: 

8. Create a graph that has at least two geom layers. 
9. Create 3 histograms, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->
## Summarizing and graphing 

### Research Question 1

**What are the 10 most commonly planted tree species in Vancouver, and what proportion of each species was planted with a root barrier?** 

*Summarizing*

To begin to answer this research question, I will compute the proportion and counts of tree species with and without a root barrier using the *summarise* and *mutate* functions.

```{r}

trees <- vancouver_trees

trees %>% group_by(genus_name, species_name, root_barrier) %>% 
  summarise(n=n()) %>% 
  mutate(freq = n/sum(n)) %>% 
   arrange(desc(n))
```
I arranged the table in descending order by count (n). I can see that the first 10 tree species are the most numerous species - answering the first part of my question, about the 10 most commonly planted tree species in Vancouver. From the freq column I learn that for the 10 most commonly planted species between 82-100% of the trees were planted without a root barrier (meaning 0-18% were planted **with** a root barrier), answering the second part of my question.


*Graphing* 

Due to the fact that there is not much variability in the proportion of trees planted with a root barrier in the top 10 species, I will instead examine the frequency distribution across species for which some trees were planted with a root barrier. I will plot three histograms with differing binwidths.

```{r}
trees %>%  group_by(genus_name, species_name, root_barrier) %>% 
  summarise(n=n()) %>% 
  arrange(desc(n)) %>% 
  mutate(freq = n/sum(n)) %>% 
  filter(root_barrier=="Y") %>% 
ggplot(aes(freq))+
  geom_histogram(binwidth=0.1)+
  labs(title="Vancouver trees planted with a root barrier", x="Proportion of trees with root barrier", y="Number of species")

```
```{r}
trees %>%  group_by(genus_name, species_name, root_barrier) %>% 
  summarise(n=n()) %>% 
  arrange(desc(n)) %>% 
  mutate(freq = n/sum(n)) %>% 
  filter(root_barrier=="Y") %>% 
ggplot(aes(freq))+
  geom_histogram(binwidth=0.2)+
  labs(title="Vancouver trees planted with a root barrier", x="Proportion of trees with root barrier", y="Number of species")
```
```{r}
trees %>%  group_by(genus_name, species_name, root_barrier) %>% 
  summarise(n=n()) %>% 
  arrange(desc(n)) %>% 
  mutate(freq = n/sum(n)) %>% 
  filter(root_barrier=="Y") %>% 
ggplot(aes(freq))+
  geom_histogram(binwidth=0.5)+
  labs(title="Vancouver trees planted with a root barrier", x="Proportion of trees with root barrier", y="Number of species")
```

The histogram with binwidth = 0.1 is the best representation of the data, because the bin sizes are small enough it retains enough detail to see the trends. The other two charts have bins that are too large and the trends are obscured.
We can see that for tree species planted with a barrier, most species had fewer than 25% planted with a barrier. 

### Research Question 2

Original research question: 


**What species of trees were planted between 1995-2000, and how do they compare to those planted between 2010-2015?** 

Modified question:

**What genera of trees were planted before 2010, and how do they compare to those planted after 2010?**

I decided to switch to a coarser scale of analysis (genera instead of species), and to change the period of my analysis. According to Vancouver's Urban Forest Strategy, in 2010, the Greenest City Action Plan set a target of planting 150,000 trees by 2020 to improve access to nature.


*Summarizing*

I can compute the number of observations for each genera using the *summarise* function.

```{r}
trees %>% group_by(genus_name) %>% 
  summarise(count=n()) %>% 
  arrange(desc(count))
```

The five most commonly planted genera are Acer, Prunus, Fraxinus, Tilia, and Quercus.  

I will use the *mutate* function to create a categorical variable called period with two groups - Pre-plan and Post-plan meaning before and after the Action Plan in 2010 - from the date_planted variable. I can then compute the number of trees in each genera planted during each period.

```{r}
trees %>% mutate(period = case_when(date_planted < '2010-01-01' ~"Pre-plan", TRUE ~ "Post-plan")) %>% 
  na.omit(date_planted) %>% 
  group_by(genus_name, period) %>% 
  summarise(count=n()) %>% 
  arrange(genus_name)
```
This table gives me a nice summary with the number of trees in each genus planted before and after 2010. 


*Graphing* 

Next I will filter by the five most common genera to plot the number of trees in each genera by planting date, to view the trends in these genera over time. I will create a jitter plot and customize the alpha transparency so that trees of different genera planted on the same dates will be visible.

```{r}
  
trees %>% filter(genus_name == c("ACER", "PRUNUS", "TILIA", "FRAXINUS","QUERCUS")) %>%
  group_by(genus_name, date_planted) %>% 
  summarise(count=n()) %>% 
na.omit(date_planted)  %>% 
ggplot(aes(date_planted, count, color=genus_name))+
  geom_jitter(alpha=0.2)+
  scale_color_discrete(name="Genus name")+
  theme_bw()+
  labs(title="Five most common tree genera by planting date", x="Date planted", y="Number of trees")

```


This exercise doesn't really answer my research question, because I am less interested in the exact date that trees were planted and more interested in trends over long periods of time. 

### Research Question 3 


**Where are the 10 tallest trees located in Vancouver, and what species are they?**

Modified research question:

**Where are the 10 largest trees located in Vancouver, and what species are they?** 

I decided to change this question from a focus on the *tallest* to *largest* trees so that I could examine the numerical variable diameter, which is tree diameter at breast height (DBH) in inches. The height_range_id variable does not give information about the actual height of a tree, but categorizes the height into 10 ft bins. 


*Summarizing*

I will use the *summarise* function to calculate summary statistics for the diameter variable by genus/species. I will calculate the mean, range, min, max, median, and spread (standard deviation).

```{r}
trees %>% group_by(genus_name, species_name) %>% 
  summarise(mean = mean(diameter), min=min(diameter), max=max(diameter), range=range(diameter), median = median(diameter), sd = sd(diameter)) %>% 
  arrange(desc(max))
```
I arranged the output in descending order by max height, which shows me the 10 largest trees by diameter. This table gives me an answer to the second part of my question. However the first three trees in the list have diameters between 305 in and 435 in - 25 to 36 ft! This seems likely to be an error (maybe circumference was measured instead of DBH). 


*Graphing* 

I will use the *filter* function to filter by max diameter (max<300) to remove these outliers, and then plot the relationship between median diameter and max diameter for the tree species in the dataset. I will use two geom layers - geom_point and geom_smooth to plot the individual data points and the fitted linear model.

```{r}
trees %>% group_by(genus_name, species_name) %>% 
  summarise(max=max(diameter), median = median(diameter)) %>% 
  arrange(desc(max)) %>% 
  filter(max<300) %>% 
  ggplot(aes(median, max))+
  geom_point()+
  geom_smooth()+
  theme_light()+
  labs(title="Median and maximum diameter by tree species", x="Median diameter (in)", y="Max diameter (in)")

```


This plot doesn't answer my original question - what neighbourhoods the largest trees are located in - but it does show some interesting relationships between median and max diameter for different species. 


### Research Question 4 


**What is the distribution of coniferous and deciduous trees by neighbourhood in Vancouver?** 


*Summarizing* 

To answer this question, I will create a new categorical variable called 'type' from an existing categorical variable - genus_name. To do this, I need to know what genera are coniferous and which are deciduous. From examining my dataset, I know that more of the trees are deciduous, therefore I identified all of the coniferous genera. I will use the *mutate* function to create a new column called type, and the *case_when* function to call out the genera that are coniferous to assign them the value 'Coniferous' and assign all other genera the value 'Deciduous.'

```{r}
trees %>% mutate(type = case_when(
  genus_name %in% c('ABIES', 'PINUS','THUJA', 'PICEA', 'PSEUDOTSUGA', 'SEQUOIADENDRON', 'CEDRUS', 'TSUGA', 'METASEQUOIA', 'ARAUCARIA', 'CRYPTOMERIA', 'CALOCEDRUS', 'TAXODIUM', 'TAXUS', 'JUNIPERUS', 'SEQUOIA', 'SCIADOPITYS', 'CUPRESSOCYPARIS X') ~'Coniferous',
     TRUE ~ 'Deciduous')) %>% 
  arrange(type)
```


*Graphing* 

Next I will group the output by neighbourhood name and type, to create a bar chart displaying the number of trees of each type by neighbourhood. I will use the function *scale_y_log10* to log transform the y axis so the differences between the number of coniferous and deciduous trees by neighbourhood can be visualized more easily.

```{r}
trees %>% mutate(type = case_when(
  genus_name %in% c('ABIES', 'PINUS','THUJA', 'PICEA', 'PSEUDOTSUGA', 'SEQUOIADENDRON', 'CEDRUS', 'TSUGA', 'METASEQUOIA', 'ARAUCARIA', 'CRYPTOMERIA', 'CALOCEDRUS', 'TAXODIUM', 'TAXUS', 'JUNIPERUS', 'SEQUOIA', 'SCIADOPITYS', 'CUPRESSOCYPARIS X') ~'Coniferous',
     TRUE ~ 'Deciduous')) %>% 
  arrange(type) %>% 
  group_by(neighbourhood_name, type) %>% 
  ggplot(aes(neighbourhood_name, fill=type))+
  scale_fill_discrete(name="Tree type")+
  geom_bar(width=1, position=position_dodge2())+
  coord_flip()+
  scale_y_log10()+
  labs(title="Distribution of tree type by neighbourhood", y="Log transformed number of trees", x="")+
  theme_minimal()
```


These two exercises enabled me to answer the research question.


<!----------------------------------------------------------------------------->

### 1.3 (2 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!------------------------- Write your answer here ---------------------------->

1. The first research question was answered based on my analysis in 1.2. It might be interesting to investigate which tree species have the highest proportion of trees planted with a root barrier, and if there is a pattern to their planting location which explains why they have a root barrier (or if it is more species-specific). 

2. The second research question has not been fully answered. The trends in trees planted before and after Vancouver's Greenest Action Plan in 2010 are still unclear. After plotting the trends in the five most common genera planted, I am now more interested in examining trends in species planted within the Acer genus, before 2010 and after 2010. 

3. The third research question has not been fully answered. It is unclear in which neighbourhoods the largest diameter tree species are planted. After investigating the data a bit more, I am interested in the trends in max and median tree diameter by neighbourhood. 

4. The fourth question was answered by my analysis. The question could be refined to investigate the proportions of different coniferous genera found in each neighbourhood.


<!----------------------------------------------------------------------------->

# Task 2: Tidy your data 

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

-   Each row is an **observation**
-   Each column is a **variable**
-   Each cell is a **value**

### 2.1 (2 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have \>8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->
**Tidy Data** 

The vancouver_trees dataset has 20 variables. I will choose the following eight variables and determine whether the data is tidy or untidy:

tree_id,
civic_number,
genus_name,
species_name,
cultivar_name,
plant_area,
on_street,
neighbourhood_name

```{r}
trees %>% select(tree_id, civic_number, genus_name, species_name, cultivar_name, plant_area, on_street, neighbourhood_name)
```

All of the columns represent a different variable. Each row is an observation for a unique tree. Each cell contains a value. Therefore, based on these eight variables, the data is tidy.

<!----------------------------------------------------------------------------->

### 2.2 (4 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".

<!--------------------------- Start your work below --------------------------->

Starting with the tidy vancouver_trees dataset, I will untidy the data by combining two variables, genus_name and species_name, into one column called "scientific name" using the *unite* function. This is untidy because each cell will contain two values. 


I will untidy the data further by using the *pivot_wider* function to create new columns with the name coming from the neighbourhood_name variable and the values from the on_street variable. 

```{r}
trees %>% select(tree_id, civic_number, genus_name, species_name, cultivar_name, plant_area, on_street, neighbourhood_name) %>% 
  unite("scientific_name", genus_name:species_name) %>% 
  pivot_wider(names_from = "neighbourhood_name", values_from = "on_street")
```
Next, I will tidy the data to it's original tidy format using the *separate* function to separate the genus and species names into their own columns, and the *pivot_longer* function to reassign the neighbourhoods to a column called neighbourhood_name and the values to a column called on_street.

```{r}
trees %>% select(tree_id, civic_number, genus_name, species_name, cultivar_name, plant_area, on_street, neighbourhood_name) %>% 
  unite("scientific_name", genus_name:species_name) %>% 
  pivot_wider(names_from = "neighbourhood_name", values_from = "on_street") %>% separate(scientific_name, c("genus_name","species_name")) %>% 
  pivot_longer(!tree_id:plant_area, names_to = "neighbourhood_name", values_to = "on_street")
```



<!----------------------------------------------------------------------------->

### 2.3 (4 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in the remaining tasks:

<!-------------------------- Start your work below ---------------------------->

1.  *What genera of trees were planted before 2010, and how do they compare to those planted after 2010?*
2.  *Where are the 10 largest trees located in Vancouver, and what species are they?*

<!----------------------------------------------------------------------------->

Explain your decision for choosing the above two research questions.

<!--------------------------- Start your work below --------------------------->

1. This question was not fully answered by my analysis. However, instead of investigating this original question, I am going to change it slightly to examine what tree species in the genus Acer were planted before 2010, and how they compare to those planted after 2010. 

2. This question was not fully answered by my previous analysis either; the investigation I have done thus far has provoked more questions. Instead of seeking to answer the rest of the original question (where the 10 largest trees are located), I will examine trends in the max and median tree diameter by neighbourhood.

<!----------------------------------------------------------------------------->

Now, try to choose a version of your data that you think will be appropriate to answer these 2 questions. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelevant columns, etc.).

(If it makes more sense, then you can make/pick two versions of your data, one for each research question.) 

<!--------------------------- Start your work below --------------------------->

*What species of trees in the genus Acer were planted before 2010, and how do they compare to those planted after 2010?* 

```{r}
trees %>% mutate(period = case_when(date_planted < '2010-01-01' ~"Pre-plan", TRUE ~ "Post-plan")) %>% 
  na.omit(date_planted) %>% 
  filter(genus_name == "ACER") %>% 
  group_by(species_name, period) %>% 
  summarise(n=n())
```


*Where are the 10 largest trees located in Vancouver, and what is the relationship between the median and max tree diameter in each neighbourhood?* 

```{r}
trees %>% group_by(neighbourhood_name, genus_name, species_name) %>% 
  summarise(max=max(diameter), median = median(diameter)) %>% 
  arrange(desc(max)) %>% 
  filter(max<300)
```


<!----------------------------------------------------------------------------->
# Task 3: Modelling

## 3.0 (no points)

Pick a research question from 1.2, and pick a variable of interest (we'll call it "Y") that's relevant to the research question. Indicate these.

<!-------------------------- Start your work below ---------------------------->

**Research Question**: *Where are the 10 largest trees located in Vancouver, and what is the relationship between the median and max tree diameter in each neighbourhood?*

**Variable of interest**: diameter

<!----------------------------------------------------------------------------->

## 3.1 (3 points)

Fit a model or run a hypothesis test that provides insight on this variable with respect to the research question. Store the model object as a variable, and print its output to screen. We'll omit having to justify your choice, because we don't expect you to know about model specifics in STAT 545.

-   **Note**: It's OK if you don't know how these models/tests work. Here are some examples of things you can do here, but the sky's the limit.

    -   You could fit a model that makes predictions on Y using another variable, by using the `lm()` function.
    -   You could test whether the mean of Y equals 0 using `t.test()`, or maybe the mean across two groups are different using `t.test()`, or maybe the mean across multiple groups are different using `anova()` (you may have to pivot your data for the latter two).
    -   You could use `lm()` to test for significance of regression coefficients.

<!-------------------------- Start your work below ---------------------------->

I will use the *lm()* function to fit a model that makes predictions on max tree diameter based on median diameter after grouping by common name and neighbourhood name.

```{r}

library(broom)

trees_dbh <- trees %>% filter(diameter <300) %>% group_by(common_name, neighbourhood_name) %>%  mutate(median=median(diameter), max=max(diameter))


dbh_lm <- lm(max ~median, data=trees_dbh)

```
<!----------------------------------------------------------------------------->

## 3.2 (3 points)

Produce something relevant from your fitted model: either predictions on Y, or a single value like a regression coefficient or a p-value.

-   Be sure to indicate in writing what you chose to produce.
-   Your code should either output a tibble (in which case you should indicate the column that contains the thing you're looking for), or the thing you're looking for itself.
-   Obtain your results using the `broom` package if possible. If your model is not compatible with the broom function you're needing, then you can obtain your results by some other means, but first indicate which broom function is not compatible.

<!-------------------------- Start your work below ---------------------------->

I will use the *glance* function from the *broom* package to assess the model fit. This function will output a one-row tibble with information related to model fit - the statistic I am looking at is the r.squared value, shown below to be 0.33 (weak positive relationship between the two variables).

```{r}

glance(dbh_lm)

```

<!----------------------------------------------------------------------------->

# Task 4: Reading and writing data

Get set up for this exercise by making a folder called `output` in the top level of your project folder / repository. You'll be saving things there.

## 4.1 (3 points)

Take a summary table that you made from Task 1, and write it as a csv file in your `output` folder. Use the `here::here()` function.

-   **Robustness criteria**: You should be able to move your Mini Project repository / project folder to some other location on your computer, or move this very Rmd file to another location within your project repository / folder, and your code should still work.
-   **Reproducibility criteria**: You should be able to delete the csv file, and remake it simply by knitting this Rmd file.

<!-------------------------- Start your work below ---------------------------->

```{r}
size_trees <- trees %>% group_by(genus_name, species_name) %>% 
  summarise(max=max(diameter), median = median(diameter)) %>% 
  arrange(desc(max)) %>% 
  filter(max<300)
  
write_csv(size_trees, here::here("trees_diameter_max_median.csv"))
```

<!----------------------------------------------------------------------------->
## 4.2 (3 points)

Write your model object from Task 3 to an R binary file (an RDS), and load it again. Be sure to save the binary file in your `output` folder. Use the functions `saveRDS()` and `readRDS()`.

-   The same robustness and reproducibility criteria as in 4.1 apply here.

<!-------------------------- Start your work below ---------------------------->

```{r}

saveRDS(dbh_lm, "trees_model.rds")

readRDS("trees_model.rds")

```


<!----------------------------------------------------------------------------->

# Overall Reproducibility/Cleanliness/Coherence Checklist 

Here are the criteria we're looking for.

## Coherence (0.5 points)

The document should read sensibly from top to bottom, with no major continuity errors. 

The README file should still satisfy the criteria from the last milestone, i.e. it has been updated to match the changes to the repository made in this milestone. 

## File and folder structure (1 points)

You should have at least three folders in the top level of your repository: one for each milestone, and one output folder. If there are any other folders, these are explained in the main README.

Each milestone document is contained in its respective folder, and nowhere else.

Every level-1 folder (that is, the ones stored in the top level, like "Milestone1" and "output") has a `README` file, explaining in a sentence or two what is in the folder, in plain language (it's enough to say something like "This folder contains the source for Milestone 1").

## Output (1 point)

All output is recent and relevant:

-   All Rmd files have been `knit`ted to their output md files. 
-   All knitted md files are viewable without errors on Github. Examples of errors: Missing plots, "Sorry about that, but we can't show files that are this big right now" messages, error messages from broken R code
-   All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
-   There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted.

Our recommendation: delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant.

## Tagged release (0.5 point)

You've tagged a release for Milestone 2. 

### Attribution

Thanks to Victor Yuan for mostly putting this together.
